Acceleration of Global Search through Dual Lipschitz Constant Estimates \thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}

### 2019-04-12 2019/NUMTA_Barkalov/LNCS/numta19.tex
Version 0

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
%\usepackage{hyperref}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates
\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Roman Strongin%\orcidID{0000-0003-0390-6695} 
\and Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
\and Semen Bevzuk%\orcidID{0000-0002-3845-5356}
}
%
\authorrunning{R. Strongin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}

The paper considers global optimization problems with a black-box objective function satisfying the Lipschitz condition. 
Efficient algorithms for this class of problems require reliable estimates of the Lipschitz constant to be introduced. 
In fundamental work \cite{Strongin2000} various approaches have been proposed to take into account both global and local properties of the objective function. 
%В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
In particular, algorithms using local estimates of the Lipschitz constant have shown their potential \cite{Sergeyev2010,Sergeyev2016}.

The new approach presented in this paper is based on simultaneous use of two estimates: one is substantially larger then the other. 
The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
Results of numerical experiments on the random sample of multidimensional functions demonstrate the efficiency of the suggested approach.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{hyperref}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}

\begin{document}
%
\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates
\thanks{This research was supported by the the Russian Science Foundation,
project No.\,16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Roman Strongin%\orcidID{0000-0003-0390-6695} 
\and Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
\and Semen Bevzuk%\orcidID{0000-0002-3845-5356}
}
%
\authorrunning{R. Strongin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper considers global optimization problems with a black-box objective 
function satisfying the Lipschitz condition. Efficient algorithms for this 
class of problems require reliable estimates of the Lipschitz constant to be 
introduced. The various approaches have been proposed to take into account both
global and local properties of the objective function. In particular, algorithms
using local estimates of the Lipschitz constant have shown their potential.
The new approach presented in this paper is based on simultaneous use of two
estimates: one is substantially larger then the other. 
The larger estimate ensures global convergence and the smaller one reduces 
the total number of trials needed to find the global optimizer.
Results of numerical experiments on the random sample of multidimensional 
functions demonstrate the efficiency of the suggested approach.

\keywords{Global optimization \and Multiextremal problems 
\and Lipschitz constant estimates}
\end{abstract}
%
%
%
\section{Introduction}

% Введение должно быть полностью новым
% Константа Липшица - основное в методах глобальной оптимизации
% Постановку задачи - дать во введении? Можно, ради экономии места
% А в журнальной статье постановку задачи вынести в отельный раздел

\section{Global search algorithm}

According to the global search algorithm, the first two trials are executed at 
the points $y^0=y(0), y^1=y(1)$. The choice of the point $y^{k+1},k\geq 1,$  
for the next $(k+1)$-th trial is defined by the following rules.

\begin{enumerate}
	\item 
	Renumber the preimages of all the points $y^i=y(x^i)$
	from the trials already performed  	
%\begin{equation}\label{y_i} 
%y^0=y(x^0), y^1=y(x^1),...,y^k=y(x^k)
%\end{equation}
by subscripts in the increasing order of their coordinates, i.e.
\begin{equation}\label{x_i}
0=x_0<x_1<\dots <x_k=1,
\end{equation}
and associate these with the values $z_i=\varphi(y(x_i)), 0\leq i \leq k,$ 
computed at these points.
\item
Compute the maximum absolute value of the first divided differences
\begin{equation}\label{mu}
\mu = \max_{1 \leq i \leq k}\frac{\left|z_i-z_{i-1}\right|}{\Delta_i},
\end{equation}
where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$. If $\mu = 0$, set $\mu = 1$.
\item
For each interval $(x_{i-1}, x_i), \; 1\leq i \leq k,$  calculate the value
\begin{equation}\label{R}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-z^*}{r\mu}
\end{equation}
called the \textit{characteristic} of the interval; the real number $r>1$ being 
the input parameter of the algorithm.
\item 
Select the interval $(x_{t-1},x_t)$ corresponding to the maximum characteristic
\begin{equation}\label{MaxR}
R(t)= \max_{1 \leq i \leq k}R(i).
\end{equation}
\item
Carry out the next trial at the point $x^{k+1}\in(x_{t-1},x_t)$ calculated using
the following formula
\begin{equation}\label{xk1}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}
\left[\frac{\left|z_t-z_{t-1}\right|}{\mu}\right]^N.
\end{equation}
\end{enumerate}

The algorithm terminates if the condition $\Delta_t < \epsilon$ is satisfied
where $t$ is from (\ref{MaxR}), and $\epsilon>0$ is the predefined accuracy.

The theory of convergence of this algorithm is provided in \cite{Strongin2000}.
%The modifications taking into account .... are given in [....].

\section{Алгоритм с двойной оценкой Lipschitz constant}

Комментарий об использовании $\mu$ как оценки константы Липшица

Ключевое условие из теорема о сходимости - $r\mu > 2^{3-1/N}L\sqrt{N+3} $

Возьмем большое значение параметра $r$ - алгоритм гарантированно сойдется к глобальному минимум, но сделает много итераций, т.к. значение константы Липшица будет переоценено.

Возьмем малое значение $r$ - алгоритм будет сходиться быстро, но может потерять свойство сходимости к глобальному экстремуму, т.к. значение константы Липшица будет недооценено.

Характеристики $R(i)$ зависят от параметра $r$, и они будут несравнимы при разных значениях $r$.

Сравнимость характеристик можно получить, если ввести нормировочные коэффициент (формула).

\section{Numerical Experiments}
